Week 7 R
=====
### Survival Analysis in R

R comes preinstalled with the `survival` package to do survival analysis.
```{r}
library(survival)
```

We can return to the `rats` dataset from our week 2 example to see how this works.
```{r}
data(rats)
```

We first have to create a survival object by combining time and the event (censored or failure/tumor).
```{r}
surv.object <- Surv(time = rats$time, event = rats$status)
```

The time are labeled by + signs for censored observations.
```{r}
surv.object
```

We then fit a survival curve to it using `survfit`. As you can see, that's not so interesting, so we can plot it to see the actual curve.
```{r}
survfit(surv.object ~ 1)
plot(survfit(surv.object ~ 1))
```

To compare two curves by group, we can replace the 1 in the formula above by the group variable, in this case, treatment.
```{r}
surv.all <- survfit(surv.object ~ rats$rx)
plot(surv.all, col = c("red","blue"), xlab = "time", ylab = "percent survival")
legend(x = "bottomleft", legend = "Red = control, Blue = drug")
```

The curves look divergent but to test the null hypothesis that they are the same, we can use the log rank test. This is done using the `survdiff` function.
```{r}
survdiff(surv.object ~ rats$rx)
```

### Problems

1. Load the National Wilms Tumor dataset (`data(nwtco)`) from the survival package.
2. Create the survey object using `rel` (relapse) as the event and `edrel` (time to relapse) as the time.
3. Fit the survival curve and plot it.
4. Compare time to relapse by the tumor pathology reported by the local institution (variable name is `instit`). Plot the survival curve by local institution tumor pathology.
5. Test the null hypothesis that the two survival curves for both pathology types are equal.