Week 7 R
=====
### Survival Analysis in R

### Problems

1. Load the National Wilms Tumor dataset (`data(nwtco)`) from the survival package.

```{r}
library(survival)
data(nwtco)
str(nwtco)
```

2. Create the survey object using `rel` (relapse) as the event and `edrel` (time to relapse) as the time.

```{r}
surv.object <- Surv(time = nwtco$edrel, event = nwtco$rel)
```


3. Fit the survival curve and plot it.

```{r}
survfit(surv.object ~ 1)
plot(survfit(surv.object ~ 1))
```

4. Compare time to relapse by the tumor pathology reported by the local institution (variable name is `instit`). Plot the survival curve by local institution tumor pathology.

```{r}
survfit(surv.object ~ nwtco$instit)
plot(survfit(surv.object ~ nwtco$instit), col = c("red","blue"), xlab = "time", ylab = "percent without relapse")
legend(x = "bottomleft", legend = "Red = Pathology #1, Blue = Pathology #2")
```

5. Test the null hypothesis that the two survival curves for both pathology types are equal.

```{r}
survdiff(surv.object ~ nwtco$instit)
```
