Homework 2
========================================================

Input separate tables for person time and risks to make it easier for later calculations

```{r}
Table.person.time <- matrix(c(20, 4, 2698, 2699), 2, 2)
rownames(Table.person.time) <- c("Vioxx", "naproxen")
colnames(Table.person.time) <- c("Heart attacks", "Person-years")
Table.person.time

Table.events <- matrix(c(20, 4, 4047-20, 4029-4), 2, 2)
rownames(Table.events) <- c("Vioxx", "naproxen")
colnames(Table.events) <- c("Heart attacks", "No heart attacks")
Table.events
```

### Question 1.
What is the absolute difference in the incidence rates of heart attacks for Vioxx versus Naproxen? Please calculate as events per 1000 person-years and round to the nearest tenth (e.g., x.x); do not write the units when entering your answer.

```{r}
Vioxx.rate <- Table.person.time[1,1]/Table.person.time[1,2]
naproxen.rate <- Table.person.time[2,1]/Table.person.time[2,2]
round(1000*(Vioxx.rate - naproxen.rate), digits = 1)
```

### Question 2.
What is the absolute difference in the risk of heart attacks for Vioxx versus Naproxen? Please report as a percentage, rounded to the nearest hundredth (e.g., .xx). Do not include the % sign.

```{r}
Vioxx.risk <- Table.events[1,1]/Table.events[1,2]
naproxen.risk <- Table.events[2,1]/Table.events[2,2]
(Vioxx.risk - naproxen.risk)*100
```

### Question 3.
What is the rate ratio comparing Vioxx with Naproxen? Please round to the nearest tenth (e.g., x.x).

```{r}
Vioxx.rate/naproxen.rate
```

### Question 4.
What is the number needed to harm? Round to the nearest whole number.

Number needed to harm is the inverse of the absolute risk difference.
```{r}
1/(Vioxx.risk - naproxen.risk)
```

### Question 5.
What percent of single-pill users were hospitalized during follow-up? (do not include the % sign)
```{r}
one.pill.risk <- 7.7/100
```


### Question 6.
What percent of three-or-more-pill users were hospitalized during follow-up? (eg. x.x, do not include the % sign)

```{r}
three.pill.risk <- 9.9/100
```

### Question 7.
Calculate the risk ratio comparing the risk of hospitalization in single-pill users versus three-or-more-pill users. Round to the nearest hundredth (e.g., .xx).

```{r}
one.pill.risk/three.pill.risk
```

### Question 8.
Calculate the odds ratio comparing the odds of hospitalization in single-pill users versus three-or-more-pill users. Round to the nearest hundredth (e.g., .xx).
```{r}
one.pill.odds <- one.pill.risk/(1-one.pill.risk)
three.pill.odds <- three.pill.risk/(1-three.pill.risk)
# Odds ratio
one.pill.odds/three.pill.odds
```

### Question 10.
The adjusted odds ratio comparing the1.0 mg tesofensine group with the placebo group for the outcome of a weight loss of 5kg or more was 31.5. Calculate the unadjusted risk ratio for this comparison (using the percentages given in the table). Round to the nearest tenth (e.g., x.x).

```{r}
Table10 <- matrix(c(42, 13, 4, 32), 2, 2)
rownames(Table10) <- c("tesofensine 1.0 mg", "placebo")
colnames(Table10) <- c("loss >= 5 kg", "loss < 5 kg")
Table10
Table10.prop <- prop.table(Table10, 1)
Table10.prop
Table10.prop[1,1] / Table10.prop[2,1]
```

Alternately we can use the `epi.2by2` function from the `epiR` package. See `?epi.2by2` to see how you have to arrange the tables you give the function.

```{r}
library(epiR)
epi.2by2(Table10, method="cohort.count")
```
