HRP 258 Week 2, R Exercise
========================================================
Risk and Rate Ratios
----

This exercise will help you learn how to calculate risk and rate ratios and differences using R or Deducer.

### Loading the Data
We will use the `rats` dataset from the `survival` package.  

If you don't already have this package for some reason, type:
~~~
install.packages("survival")
~~~

To load the dataset type the following:

```{r}
library(survival)
data(rats)
```

To learn more about the dataset, type `?rats`.  

We will be interested in only 3 variables: `rx`, `time`, and `status`. As you can see in the help page for the dataset, `rx` is treatment (1=drug, 0=control), `time` is time to tumor or last follow-up, and `status` is event status (1=tumor, 0=censored). In this case, censored means that there was no tumor the last time they looked at that rat.

### Questions
Q1. Convert `rx` and `status` into factor variables. If you can, also label the levels with the appropriate names (i.e., drug, control, tumor, censored).
```{r}
rats$status.factor <- factor(rats$status, levels = 1:0, labels = c("tumor", "censored"))
rats$rx.factor <- factor(rats$rx, levels = 1:0, labels = c("drug", "control"))
```

Q2. Create a 2x2 contingency table with `rx` on the rows and `status` on the columns.

```{r}
Table1 <- with(rats, table(rx.factor, status.factor))
Table1
```

Q3. Using this table, what is the risk of tumor in the drug (`rx` = 1) group? What is the risk of tumor in the control (`rx` = 0) group?

```{r}
Table1risk <- prop.table(Table1, margin = 1)
Table1risk
```

Q4. Using this table, what is the risk ratio of tumor in the drug group versus the control group? What is the odds ratio? What is the risk difference? 
```{r}
Table1risk[1, 1] / Table1risk[2, 1] # Risk ratio
(Table1[1, 1] / Table1[1, 2]) / (Table1[2, 1]/ Table1[2, 2]) # Odds ratio
Table1risk[1, 1] - Table1risk[2, 1] # Risk difference
```

Q5. How much person-time (rat-time?) was observed in rats receiving the drug (`rx` = 1)? How much person-time was observed in rats receiving the control (`rx` = 0)? There are many ways to do this but one way would be to subset the data on `rx` = 1 and see what is the sum of `time`. Then re-subset on `rx` = 0 and see how much person-time there is.

```{r}
drug.time <- with(rats, sum(time[rx == 1])) # drug group
drug.time
control.time <- with(rats, sum(time[rx == 0])) # control group
control.time
```

Another solution is to use `tapply` to subgroup the data.

```{r}
person.time <- tapply(rats$time, rats$rx.factor, sum)
person.time
```


Q6. Create a 2x2 table with the rows being the `rx` groups and the first column being the number of tumors in each group and the second column being person time in each group.
```{r}
Table2 <- matrix(c(Table1[, 1], drug.time, control.time), nrow = 2, ncol = 2)
colnames(Table2) <- c("tumors", "person-time")
rownames(Table2) <- c("drug", "control")
Table2
```


Q7. What is the rate ratio comparing tumor incidence in the drug and control groups? What is the rate difference? What is the number needed to harm?
```{r}
rate.drug <- Table2[1, 1]/Table2[1, 2] 
rate.control <- Table2[2, 1] / Table2[2, 2]
rate.drug / rate.control # rate ratio
rate.drug - rate.control # rate difference
1 / (rate.drug - rate.control) # NNT
```

An alternate way to do this is to use the `epiR` package. This package has a function called `epi.2by2` that automates much of the calculations above. The only issue is that it requires you to give it tables in a specific orientation (exposures or treatments on the rows and disease on the columns). For more information, see the `epiR` documentation. To install it, type `install.packages("epiR")` at the console.

```{r}
library(epiR)
epi.2by2(Table1, method = "cohort.count")
epi.2by2(Table2, method = "cohort.time")
1 / epi.2by2(Table2, method = "cohort.time", verbose = T)$AR * 100
```
