Homework 4
========================================================

## Question 1:

### The following table shows a discrete probability distribution.
x | p(x)
---|---
0|  0
1|	.1
2|	.2
3|	.4
4|	.1
5|	.2

### Find the expected value (mean) of this probability distribution. Please round to 1 decimal place.

```{r}
x <- 0:5
px <- c(0, 0.1, 0.2, 0.4, 0.1, 0.2)
ev.x <- sum(x*px)
ev.x
```

## Question 2:

### The following table shows a discrete probability distribution.
x | p(x)
---|---
0|  0
1|	.1
2|	.2
3|	.4
4|	.1
5|	.2

### Find the standard deviation. Please round to 2 decimal places. 

```{r}
varx <- sum((x-ev.x)^2*px)
varx^0.5
```

## Question 3:

### If spine bone density is normally distributed in young women with a mean of 1.0 g/cm2 and a standard deviation of .10 g/cm2, then how many young women out of 100 would you expect to have bone densities less than .85 g/cm2?

### Round to the nearest whole number.

```{r}
mean3 <- 1
sd3 <- 0.1
100*pnorm(0.85, mean = mean3, sd = sd3)

```
