Homework 5
========================================================

Question 2:

In a study of 13 children with cystic fibrosis who completed an exercise program, exercise endurance (duration on an exercise test) improved by a mean of 0.77 minutes. The standard deviation for the change in endurance was 0.83 minutes. What is the theoretical distribution of the mean?

```{r}
sem2 <- 0.83/(13)^.5
sem2
```

Question 3:

Thirty heart disease patients are put on an exercise regimen. Twenty patients improve on the exercise stress test and ten don’t improve or get worse. Calculate the 95% confidence interval for the true proportion of heart disease patients who improve their fitness using this particular exercise regimen. Recall that proportions are normally distributed with a standard error of

```{r}
library(binom)
binom.confint(20, 30, conf.level = 0.95, methods = "asymptotic")
```

Question 4:

The following data were collected from a case-control study of breast cancer and fat intake:

 | case  | control
--- | --- | ---
High-fat diet	| 10	| 5
Low-fat diet	| 40	| 55
 | 50 |	60

Statistical inferences for odds ratios are based on the natural log of the odds ratio, rather than the odds ratio itself (because the distribution for an odds ratio does not follow a normal distribution). The sampling distribution of the natural log of the odds ratio (lnOR) follows a normal distribution, with standard error

=1a+1b+1c+1d‾‾‾‾‾‾‾‾‾‾‾‾‾‾‾‾√
(where a, b, c, and d are the cells in the 2x2 table).

Calculate the odds ratio for breast cancer (comparing high-fat diet to low-fat diet) from the 2x2 table above.

```{r}
Table4 <- matrix(c(10, 40, 5, 55), nrow = 2)
rownames(Table4) <- c("High-fat diet", "Low-fat diet")
colnames(Table4) <- c("case", "control")
Table4
```

OR
```{r}
10*55/(5*40)
```

```{r}
library(epiR)
epi.2by2(Table4, method = "case.control")
```
