Week 5 R
====

Simulate data for case control study. Prevalence of exposure in cases is 0.2, Prevalence of exposure in controls is 0.1. 

```{r}
set.seed(8)
disease <- c(rep(1, 1000), rep(0, 1000))
exposure <- c(rbinom(1000, 1, 0.2), rbinom(1000, 1, 0.1))
```

Reorder factor (not actually necessary here since the OR will be the same)

```{r}
disease <- factor(disease, levels = 1:0, labels = c("case", "control"))
exposure <- factor(exposure, levels = 1:0, labels = c("exposed", "not exposed"))
Table1 <- table(exposure, disease)
Table1
```

What is the odds ratio?

```{r}
a <- Table1[1, 1]
b <- Table1[1, 2]
c <- Table1[2, 1]
d <- Table1[2, 2]
OR1 <- (a/b) / (c/d)  # OR
OR1
```

What is the standard error of the natural log of the odds ratio?

```{r}
seOR <- sqrt(1/a + 1/b + 1/c + 1/d)
seOR
```

What is the 95% confidence interval for the natural log of the odds ratio?

```{r}
log(OR1) + qnorm(0.025)*seOR
log(OR1) + qnorm(0.975)*seOR
```

What is the 95% confidence interval for the odds ratio?

```{r}
exp(log(OR1) + qnorm(0.025)*seOR)
exp(log(OR1) + qnorm(0.975)*seOR)
```

There are a variety of functions that can help with this calculation.

The package `epitools` has an `oddsratio` function that allows you to use different methods to calculate the odds ratio and 95% confidence interval. The `method = "wald"` argument gives you the method taught in class.
```{r}
library(epitools)
oddsratio(Table1, method = "wald")
```

The `fisher.test` function from the `stats` package calculates the odds ratio using another method.

```{r}
fisher.test(Table1)
```

The `epi.2by2` function from the `epiR` package uses `fisher.test` to calculate the odds ratio and also supplies a variety of other useful estimates.
```{r}
library(epiR)
epi.2by2(Table1, method = "case.control")
```
