Week 5 R
====

We can simulate data for a case control study to practice calculating the odds ratio. Let's assume the prevalence of exposure in cases is 0.2, and the prevalence of exposure in controls is 0.1. We can generate some pseudo-random data using the code below.

```{r}
set.seed(706)
disease <- c(rep(1, 1000), rep(0, 1000))
exposure <- c(rbinom(1000, 1, 0.2), rbinom(1000, 1, 0.1))
```

Reorder factor (not actually necessary here since the OR will be the same)

```{r}
disease <- factor(disease, levels = 1:0, labels = c("case", "control"))
exposure <- factor(exposure, levels = 1:0, labels = c("exposed", "not exposed"))
Table1 <- table(exposure, disease)
Table1
```

1. What is the odds ratio?
2. What is the standard error of the natural log of the odds ratio?
3. What is the 95% confidence interval for the natural log of the odds ratio?
4. What is the 95% confidence interval for the odds ratio?
5. Install the package `epitools` and use its `oddsratio` function to calculate the odds ratio. Is it the same? Why or why not?

To set this up, start with:
```{r}
install.packages("epitools")
library(epitools)
?oddsratio
```
